\tool: Programmable, Scalable, Efficient Secure Two-Party Computation

### 2017-10-23 OaklandSubmission/ld.tex
pass over s3 fixing various nits

In this section we prove the correctness and security of our compiler.
%
We first formalize the source and target languages. Our source runtime
semantics is a model of the ideal, trusted third-party semantics, and
generates observations corresponding to the values revealed to the
parties.
%
The target language semantics (a model of the C++ code generated by
our compiler implementation) ``computes away'' the public parts and
arrays from the compiled program, generating a secure computation
circuit.
%
Finally, we formalize the circuit semantics that computes the
generated circuit, and like the source semantics, emits observations.

We then present the compilation judgments. To prove the correctness
of our compiler, we prove that it preserves the observations. For
security of our compiler, we reduce the security argument to the
security of the cryptographic protocol used to compute the secure
computation circuit.

@@

    \inferrule*[lab={\footnotesize{E-LoopI}}]
               {
                 \rho(x) \leq n\\\\
                 \rho \vdash s \Downarrow \rho_{1}; O_{1}\\\\
                 \rho_{2} = [\rho_{1}]_{\mathsf{dom}(\rho)}[x \mapsto \rho_{1}(x) + 1]\\\\
                 \rho_{2} \vdash \loops{x}{n}{s} \Downarrow \rho'; O_{2}
               }
               {
                 \rho \vdash \loops{x}{n}{s} \Downarrow \rho'; O_{1}, O_{2}
               }

@@

    \inferrule*[lab={\footnotesize{E-If}}]
               {
                 \rho \vdash e \Downarrow c\\\\
                 c = \top \Rightarrow s = s_{1}\\\\
                 c = \bot \Rightarrow s = s_{2}\\\\
                 \rho \vdash s \Downarrow \rho'; O
               }
               {
                 \rho \vdash \ite{e}{s_{1}}{s_{2}} \Downarrow \rho'; O
               }

@@

\subsubsection*{Source language}Our source language
(Figure~\ref{fig:srclang}) is a simple imperative language. Types
$\psi$ in the language consist of the base types $\sigma$, and arrays
of base types $\sigma[n]$, where $n$ is the array length. Though we
model only one dimensional arrays, our implementation supports higher
dimensional arrays as well. Expressions in the language include the
integer constants $n$, \kw{bool} constants $\top$ and $\bot$,
variables $x$, binary operations $e_{1} + e_{2}$ and $e_{1} > e_{2}$,
conditionals $\cond{e}{e_{1}}{e_{2}}$, array literals 
$[\overline{e_{i}}]_{n}$\footnote{We write $\overline{e}$ (and
  similarly for other symbols) to denote a sequence of expressions.
The length of the sequence is usually clear from the context.}, and
array reads $x[e]$. The expression $\kw{in}_{j}$ denotes input from
party $j$. The statements $s$ in the language comprise of variable
declarations and assignments ($\psi\:x = e$ and $x := e$ resp.),
\kw{for} loops, array writes ($x[e_{1}] := e_{2}$), \kw{if}
statements, and sequence of statements ($s_{1}; s_{2}$). The statement
$\kw{out}\:e$ denotes revealing the value of $e$ to the
parties. The \kw{while} statement is an internal syntax that is not
exposed to the programmer.

@@

\vspace{0.2cm}
Values consist of constants and array values. Runtime environment
$\rho$ maps variables to values. Observations are sequences of
constants.

The judgment $\rho \vdash e \Downarrow v$ denotes the big-step
evaluation of an expression~$e$ to a value~$v$ under the runtime
environment~$\rho$. Rule ({\sc{E-Var}}) looks up the value of $x$ in
the environment. Rule ({\sc{E-Add}}) inductively evaluates $e_{1}$ and
$e_{2}$ and evaluates to their addition. Rule ({\sc{E-Read}})
evaluates an array read operation. It first evaluates $x$ to an array
value $[\overline{c_{i}}]_{n_{1}}$, and $e$ to an \kw{uint} value
$n$. It then returns $c_{n}$, the $n$-th index value in the array,
provided $n < n_{1}$, the length of the array. Rule ({\sc{E-Inp}})
evaluates to some constant $c$ denoting party $j$'s input. We model
the inputs to be base constants, an array input can be written in the
language as $[\kw{in}_{j}]_{n}$, which can then evaluate using the
rule ({\sc{E-Arr}}). The remaining rules are straightforward, and are
elided for space reasons.

The judgment $\rho \vdash s \Downarrow \rho'; O$ represents the
big-step evaluation of an statement $s$ under environment $\rho$
producing a new environment $\rho'$ and observations $O$. Rule
({\sc{E-Decl}}) evaluates the expression $e$ to $v$, and returns the
updated environment $\rho[x \mapsto v]$, with empty observations. Rule
({\sc{E-If}}) evaluates the guard expression, and then evaluates
either $s_{1}$ or $s_{2}$ accordingly. $\kw{for}$ statements evaluate
through the internal $\kw{while}$ syntax. Specifically, rule
({\sc{E-For}}) updates $\rho$ with the initial counter value $n_{1}$,
evaluates $\loops{x}{n_{2}}{s}$ to $\rho_{1}; O$, and returns
$\rho_{1} - \{x\}$ (removing $x$ from the environment) and $O$. Rule
({\sc{E-LoopI}}) shows the inductive case for \kw{while}
statements, when $\rho(x) \leq n$. The rule evaluates $s$, producing
$\rho_{1}; O_{1}$. It then restricts $\rho_{1}$ to the domain of
$\rho$ ($[\rho_{1}]_{\mathsf{dom}(\rho)}$) to remove the variables
added by $s$, increments the value if $x$, and evaluates the
\kw{while} statement under this updated environment. Rule
({\sc{E-LoopT}}) is the termination case for \kw{while} statements,
when $\rho(x) > n$. Finally, the rule ({\sc{E-Out}}) evaluates the
expression, and adds it value to the observations.

@@

    \inferrule*[lab={\footnotesize{S-Inp}}]
               {
               }
               {
                 \widetilde{\rho} \vdash \kw{in}^{m}_{j} \Downarrow \kw{in}^{m}_{j}
               }
  \end{array}
  \]
  \\\\
    \fbox{$\widetilde{\rho} \vdash \widetilde{c} \Downarrow \widetilde{\rho'}; \chi$}
  \[
  \\
  \begin{array}{c}
    \inferrule*[lab={\footnotesize{S-DeclC}}]
               {
                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \widetilde{v}\\\\
                 \widetilde{v} = c \vee \widetilde{v} = [\overline{c_{i}}]_{n}\\\\
                 \widetilde{\rho'} = \widetilde{\rho}, x \mapsto \widetilde{v}
               }
               {
                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho'}; \cdot
               }

@@

    \inferrule*[lab={\footnotesize{S-DeclCkt}}]
               {
                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \kappa\quad
                 \mathsf{fresh}\:r\\\\
                 \widetilde{\rho'} = \widetilde{\rho}, x \mapsto r\quad
                 \chi = \kw{bind}\:\kappa\:r
               }
               {
                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho'}; \chi
               }

\\\\
    \inferrule*[lab={\footnotesize{S-DeclCktA}}]
               {
                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow [\overline{\kappa_{i}}]_{n}\\\\
                 \forall i \in \{0 \dots n - 1\}.\:\mathsf{fresh}\:r_{i}\\\\
                 \widetilde{\rho'} = \widetilde{\rho}, x \mapsto [r_{i}]_{n}\quad
                 \chi = \overline{\kw{bind}\:\kappa_{i}\:r_{i}}
               }
               {
                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho'}; \chi
               }

@@

               \\\\
    \inferrule*[lab={\footnotesize{S-If}}]
               {
                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow c\\\\
                 c = \top \Rightarrow \widetilde{s} = \widetilde{s_{1}}\\\\
                 c = \bot \Rightarrow \widetilde{s} = \widetilde{s_{2}}\\\\
                 \widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho'}; \chi
               }
               {
                 \widetilde{\rho} \vdash \ite{\widetilde{e}}{\widetilde{s_{1}}}{\widetilde{s_{2}}} \Downarrow \widetilde{\rho'}; \chi
               }

@@

    \inferrule*[lab={\footnotesize{S-WriteCkt}}]
               {
                 \widetilde{\rho} \vdash x \Downarrow [\overline{r_{i}}]_{n}\quad
                 \widetilde{\rho} \vdash \widetilde{e_{1}} \Downarrow n_{1}\\\\
                 n_{1} < n\quad
                 \mathsf{fresh}\:r\quad
                 \widetilde{\rho} \vdash \widetilde{e_{2}} \Downarrow \kappa\\\\
                 \widetilde{\rho'} = \widetilde{\rho}[x \mapsto [\overline{r_{i}}]_{n}[n_{1} \mapsto r]]
               }
               {
                 \widetilde{\rho} \vdash x[\widetilde{e_{1}}] := \widetilde{e_{2}} \Downarrow \widetilde{\rho'}; \kw{bind}\:\kappa\:r
               }

@@

\subsubsection*{Target language} Figure~\ref{fig:tgtlang} shows the
target language of our compiler. The syntax follows that of the source
languag, except that the types and operators are \emph{labeled}. Below
we mainly focus on the bits that are different from the source
language.

@@

The target semantics models the semantics of the code output by
our compiler (C++ code in our implementation). It computes over the
public parts of the program, emitting a circuit that can later be
evaluated using a cryptographic MPC protocol. The semantics also
flattens the circuit so that the circuits are unaware of the array
structure. Crucially, this phase of the semantics does not have access
to the secrets.

Figure~\ref{fig:tgtruntime} shows the runtime syntax. A wire id range
$r$ denotes a set of circuit wires that carry the value of a secret
shared variable in the output circuit. Circuit gates $\kappa$ are
$r$, input gates, \kw{add}, \kw{gt}, and \kw{mux} gates, and coerce
gates $\widetilde{w} \rhd m$. Target values $\widetilde{v}$ then
consist of base values $c$ and $\kappa$, and arrays of base values.

Figure~\ref{fig:tgtsem} shows the judgments for the target
semantics. We first focus on the expression evaluation judgment
$\widetilde{\rho} \vdash \widetilde{e} \Downarrow
\widetilde{v}$. Rules ({\sc{S-PAdd}}) and ({\sc{S-SAdd}}) illustrate
the significance of the operator labels. In particular, the rule
({\sc{S-PAdd}}) evaluates a public addition $\widetilde{e_{1}}
+_{\mathcal{P}} \widetilde{e_{2}}$ to $n_{1} + n_{2}$, similar to the
source language. In contrast, the rule
({\sc{S-SAdd}}) evaluates a secret addition $\widetilde{e_{1}}
+_{\mathcal{A}} \widetilde{e_{2}}$, using an arithmetic add
gate, to $\kw{add}\:\kappa_{1}\:\kappa_{2}$. Since our compiler
compiles secret add operations to arithmetic gates, the target
expressions from our compiler never have $\widetilde{e_{1}}
+_{\mathcal{B}} \widetilde{e_{2}}$ (we present the compilation rules
later in the section). Rules ({\sc{S-PCond}}) and ({\sc{S-SCond}})
are along the similar lines. Rule ({\sc{S-PCond}}) evaluates a public
conditional to the values from one of the branches, while the rule
({\sc{S-SCond}}) evaluates to a \kw{mux} gate that takes input
circuits from the conditional ($\kappa_{1}$) and both the branches
($\kappa_{2}$ and $\kappa_{3}$). Rules ({\sc{S-Coerce}}) and
({\sc{S-Inp}}) evaluate to the coerce and input gates respectively.

Statement evaluation $\widetilde{\rho} \vdash \widetilde{s} \Downarrow
\widetilde{\rho'}; \chi$ evaluates a target statement to produce a new
environment $\widetilde{\rho'}$, and a circuit $\chi$. $\chi$ (shown
in Figure~\ref{fig:tgtlang}) is either empty, \kw{bind}-ing of a
circuit gate $\kappa$ to range $r$, \kw{out} gate, or a sequence of
circuits. Rules ({\sc{S-DeclC}}), ({\sc{S-DeclCkt}}), and
({\sc{S-DeclCktArr}}) show the variable declaration cases. Rule
({\sc{S-DeclC}}) shows the case when $\widetilde{e}$ evaluates to
$\widetilde{v}$, where $\widetilde{v}$ is a constant or an array of
constants. In this case, the mapping $x \mapsto \widetilde{v}$ is
added to the environment, and the resulting circuit is empty. When
$\widetilde{e}$ evaluates to a circuit $\kappa$, rule
({\sc{S-DeclCkt}}) picks a fresh $r$, adds the mapping $x \mapsto r$
to the environment, and outputs the circuit
$\kw{bind}\:\kappa\:r$. Rule ({\sc{S-DeclCktA}}) is analogous for
$\widetilde{e}$ evaluating to an array of circuits. The variable
assignment rules (not shown in the figure) are similar. Rule
({\sc{S-WriteCkt}}) shows the case for writing to an array, where the
array contents are secret. Finally, rule ({\sc{S-Out}}) compiles to an
\kw{out} circuit.

@@

During the circuit evaluation, the wire ranges $r$ are
mapped to (opaque) strings $b$. The semantics of these strings is
given by pairs of encryption-decryption functions, written as
$\mathcal{E}_{m}$ and $\mathcal{D}_{m}$ (where $m$ is either
$\mathcal{A}$ or $\mathcal{B}$). More concretely,
$\mathcal{E}_{m}(c)$ returns a pair of two string $b_{1}$ and $b_{2}$
(shares of the two parties), with the property that
$\mathcal{D}_{m}(b_{1}, b_{2}) = c$. We assume that the underlying MPC
protocol instantiates $\mathcal{E}_{m}$ and $\mathcal{D}_{m}$
appropriately.

The circuit semantics is shown in Figure~\ref{fig:cktsem} using the
judgments $\widehat{\rho_{1}}, \widehat{\rho_{1}} \vdash \kappa
\Downarrow b_{1}, b_{2}$, and $\widehat{\rho_{1}}, \widehat{\rho_{2}}
\vdash \chi \Downarrow \widehat{\rho_{1}'}, \widehat{\rho_{2}'}; O$,
where $\widehat{\rho_{1}}$ and $\widehat{\rho_{2}}$ are the circuit
environments of the two parties, mapping wire ranges $r$ to strings
$b$. Rule {\sc{C-Add}} evinces the pattern for evaluating circuit
gates $\kappa$. To evaluate $\kw{add}\:\kappa_{1}\:\kappa_{2}$, the
rule first evaluates $\kappa_{1}$ to $(b_{11}, b_{21})$ and
$\kappa_{2}$ to $(b_{12}, b_{22})$. Shares $(b_{11}, b_{21})$ are then
combined using $\mathcal{D_{\mathcal{A}}}$ to $n_{1}$, and similarly
$(b_{12}, b_{22})$ are combined to $n_{2}$. The final output of the
\kw{add} gate is then $\mathcal{E}_{\mathcal{A}}(n_{1} + n_{2})$. Note
that this is a functional description of how the \kw{add} gate
evaluates, of course, concretely $n_{1}$ and $n_{2}$ are not observed
by the parties. Rule {\sc{C-Coerce}} re-encrypts the shares using the
scheme for $m$. Coming to the evaluation of circuits, the evaluation
of \kw{bind} updates the mapping of $r$ in the input environments, and
the rule {\sc{C-Out}} outputs the clear value $c$ to the observations.

@@

     \inferrule*[lab={\footnotesize{T-Cond}}]
               {
                 \Gamma \vdash e : \kw{bool}^{\ell} \leadsto \widetilde{e}\\\\
                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \sigma^{\ell'} \leadsto \widetilde{e_{i}}\\\\
                 \ell = \mathcal{P} \vee (\ell = \mathcal{B} \wedge \ell' =\mathcal{B})
               }
               {
                 \Gamma \vdash \cond{e}{e_{1}}{e_{2}} : \sigma^{\ell'} \leadsto \lcond{\ell}{\widetilde{e}}{\widetilde{e_{1}}}{\widetilde{e_{2}}}
               }

@@

  \end{array}
  \]
  \\\\
    \fbox{$\Gamma \vdash s \leadsto \widetilde{s} \mid \Gamma'$}
  \[
  \\
  \begin{array}{c}
     \inferrule*[lab={\footnotesize{T-Decl}}]
               {
                 \psi = \sigma \Rightarrow \tau = \sigma^{\ell}\\\\
                 \psi = \sigma[n] \Rightarrow \tau = \sigma^{\ell}[n]\\\\
                 \Gamma \vdash e : \tau \leadsto \widetilde{e}
               }
               {
                 \Gamma \vdash \psi\:x = e \leadsto \tau\:x = \widetilde{e} \mid \Gamma, x:\tau
               }

@@

     \inferrule*[lab={\footnotesize{T-For}}]
               {
                 \Gamma' = \Gamma, x:\kw{uint}^{\mathcal{P}}\\\\
                 \Gamma' \vdash \loops{x}{n_{2}}{s} \leadsto \loops{x}{n_{2}}{\widetilde{s}} \mid \Gamma'
               }
               {
                 \Gamma \vdash \for{x}{n_{1}}{n_{2}}{s} \leadsto \for{x}{n_{1}}{n_{2}}{\widetilde{s}} \mid \Gamma
               }

@@

\subsubsection*{Compilation judgments} We now come to the compilation
rules shown in Figure~\ref{fig:compile}. We present the rules in a
declarative style, where the rules are also non-syntax
directed, and the labels $\ell$ are also chosen
non-deterministically.

>>>>

In this section we prove the correctness and security of our compiler.
%
We first formalize the source and target languages. Our source runtime
semantics is a model of the ideal, trusted third-party semantics, and
generates observations corresponding to the values revealed to the
parties.
%
The target language semantics (a model of the C++ code generated by
our compiler implementation) ``computes away'' the public parts and
arrays from the compiled program, generating a secure computation
circuit. Crucially, this semantics does not have access to the secret
inputs--those are processed by the secure computation circuit.
%
Finally, we formalize the circuit semantics that computes the
generated circuit, and like the source semantics, emits observations
corresponding to the values revealed to the parties.

We then present the compilation rules. To prove the correctness
of our compiler, we prove that it preserves the observations. For
security of our compiler, we reduce the security argument to the
security of the cryptographic protocol used to compute the secure
computation circuit.

@@

    \inferrule*[lab={\footnotesize{E-LoopI}}]
               {
                 \rho(x) \leq n\\\\
                 \rho \vdash s \Downarrow \rho_{1}; O_{1}\\\\
                 \rho_{2} = [\rho_{1}]_{\mathsf{dom}(\rho)}[x \mapsto \rho_{1}(x) + 1]\\\\
                 \rho_{2} \vdash \loops{x}{n}{s} \Downarrow \rho_{3}; O_{2}
               }
               {
                 \rho \vdash \loops{x}{n}{s} \Downarrow \rho_{3}; O_{1}, O_{2}
               }

@@

    \inferrule*[lab={\footnotesize{E-If}}]
               {
                 \rho \vdash e \Downarrow c\\\\
                 c = \top \Rightarrow s = s_{1}\\\\
                 c = \bot \Rightarrow s = s_{2}\\\\
                 \rho \vdash s \Downarrow \rho_{1}; O
               }
               {
                 \rho \vdash \ite{e}{s_{1}}{s_{2}} \Downarrow \rho_{1}; O
               }

@@

\subsubsection*{Source language}Our source language
(Figure~\ref{fig:srclang}) is a simple imperative language. Types
$\psi$ consist of the base types $\sigma$, and arrays
of base types $\sigma[n]$, where $n$ is the array length. Though we
model only one dimensional arrays, our implementation supports higher
dimensional arrays as well. Expressions in the language include the
integer constants $n$, \kw{bool} constants $\top$ and $\bot$,
variables $x$, binary operations $e_{1} + e_{2}$ and $e_{1} > e_{2}$,
conditionals $\cond{e}{e_{1}}{e_{2}}$, array literals 
$[\overline{e_{i}}]_{n}$\footnote{We write $\overline{e}$ (and
  similarly for other symbols) to denote a sequence of expressions.
The length of the sequence is usually clear from the context.}, and
array reads $x[e]$. The expression $\kw{in}_{j}$ denotes input from
party $j$. The statements $s$ in the language comprise of variable
declarations and assignments ($\psi\:x = e$ and $x := e$ resp.),
\kw{for} loops, array writes ($x[e_{1}] := e_{2}$), \kw{if}
statements, and sequence of statements ($s_{1}; s_{2}$). The statement
$\kw{out}\:e$ denotes revealing the value of $e$ to the
parties. The \kw{while} statement is an internal syntax that is not
exposed to the programmer.

@@

\vspace{0.2cm}
Values consist of constants and array values, runtime environment
$\rho$ maps variables to values, and observations are sequences of
constants.

The judgment $\rho \vdash e \Downarrow v$ denotes the big-step
evaluation of an expression~$e$ to a value~$v$ under the runtime
environment~$\rho$. Rule ({\sc{E-Var}}) looks up the value of $x$ in
the environment. Rule ({\sc{E-Add}}) inductively evaluates $e_{1}$ and
$e_{2}$, and evaluates to their sum. Rule ({\sc{E-Read}})
evaluates an array read operation. It first evaluates $x$ to an array
value $[\overline{c_{i}}]_{n_{1}}$, and $e$ to a \kw{uint} value
$n$. It then returns $c_{n}$, the $n$-th index value in the array,
provided $n < n_{1}$, the length of the array. Rule ({\sc{E-Inp}})
evaluates to some constant $c$ denoting party $j$'s input. We model
the inputs to be base constants, an array input can be written in the
language as $[\kw{in}_{j}]_{n}$, which can then evaluate using the
rule ({\sc{E-Arr}}). The remaining rules are straightforward, and are
elided for space reasons.

The judgment $\rho \vdash s \Downarrow \rho_{1}; O$ represents the
big-step evaluation of a statement $s$ under environment $\rho$
producing a new environment $\rho_{1}$ and observations $O$. Rule
({\sc{E-Decl}}) evaluates the expression $e$ to $v$, and returns the
updated environment $\rho[x \mapsto v]$, with empty observations. Rule
({\sc{E-If}}) evaluates the guard expression, and then evaluates
either $s_{1}$ or $s_{2}$ accordingly. $\kw{for}$ statements evaluate
through the internal $\kw{while}$ syntax. Specifically, the rule
({\sc{E-For}}) updates $\rho$ with the initial counter value $n_{1}$,
evaluates $\loops{x}{n_{2}}{s}$ to $\rho_{1}; O$, and returns
$\rho_{1} - \{x\}$ (removing $x$ from $\rho_{1}$) and $O$. Rule
({\sc{E-LoopI}}) shows the inductive case for \kw{while}
statements, when $\rho(x) \leq n$. The rule evaluates $s$, producing
$\rho_{1}; O_{1}$. It then restricts $\rho_{1}$ to the domain of
$\rho$ ($[\rho_{1}]_{\mathsf{dom}(\rho)}$) to remove the variables
added by $s$, increments the value if $x$, and evaluates the
\kw{while} statement under this updated environment. Rule
({\sc{E-LoopT}}) is the termination case for \kw{while} statements,
when $\rho(x) > n$. Finally, the rule ({\sc{E-Out}}) evaluates the
expression, and adds its value to the observations.

@@

    \inferrule*[lab={\footnotesize{S-Inp}}]
               {
               }
               {
                 \widetilde{\rho} \vdash \kw{in}^{m}_{j} \Downarrow \kw{in}^{m}_{j}
               }
  \end{array}
  \]
  \\\\
    \fbox{$\widetilde{\rho} \vdash \widetilde{c} \Downarrow \widetilde{\rho'}; \chi$}
  \[
  \\
  \begin{array}{c}
    \inferrule*[lab={\footnotesize{S-DeclC}}]
               {
                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \widetilde{v}\\\\
                 \widetilde{v} = c \vee \widetilde{v} = [\overline{c_{i}}]_{n}\\\\
                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto \widetilde{v}
               }
               {
                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}_{1}; \cdot
               }

@@

    \inferrule*[lab={\footnotesize{S-DeclCkt}}]
               {
                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \kappa\quad
                 \mathsf{fresh}\:r\\\\
                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto r\quad
                 \chi = \kw{bind}\:\kappa\:r
               }
               {
                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}_{1}; \chi
               }

\\\\
    \inferrule*[lab={\footnotesize{S-DeclCktA}}]
               {
                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow [\overline{\kappa_{i}}]_{n}\\\\
                 \forall i \in \{0 \dots n - 1\}.\:\mathsf{fresh}\:r_{i}\\\\
                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto [r_{i}]_{n}\quad
                 \chi = \overline{\kw{bind}\:\kappa_{i}\:r_{i}}
               }
               {
                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}_{1}; \chi
               }

@@

               \\\\
    \inferrule*[lab={\footnotesize{S-If}}]
               {
                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow c\\\\
                 c = \top \Rightarrow \widetilde{s} = \widetilde{s_{1}}\\\\
                 c = \bot \Rightarrow \widetilde{s} = \widetilde{s_{2}}\\\\
                 \widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \chi
               }
               {
                 \widetilde{\rho} \vdash \ite{\widetilde{e}}{\widetilde{s_{1}}}{\widetilde{s_{2}}} \Downarrow \widetilde{\rho}_{1}; \chi
               }

@@

    \inferrule*[lab={\footnotesize{S-WriteCkt}}]
               {
                 \widetilde{\rho} \vdash x \Downarrow [\overline{r_{i}}]_{n}\quad
                 \widetilde{\rho} \vdash \widetilde{e_{1}} \Downarrow n_{1}\\\\
                 n_{1} < n\quad
                 \mathsf{fresh}\:r\quad
                 \widetilde{\rho} \vdash \widetilde{e_{2}} \Downarrow \kappa\\\\
                 \widetilde{\rho}_{1} = \widetilde{\rho}[x \mapsto [\overline{r_{i}}]_{n}[n_{1} \mapsto r]]
               }
               {
                 \widetilde{\rho} \vdash x[\widetilde{e_{1}}] := \widetilde{e_{2}} \Downarrow \widetilde{\rho}_{1}; \kw{bind}\:\kappa\:r
               }

@@

\subsubsection*{Target language} Figure~\ref{fig:tgtlang} shows the
target language of our compiler. The syntax follows that of the source
language, except that the types and operators are \emph{labeled}. Below
we mainly focus on the bits that are different from the source
language.

@@

The target semantics models the semantics of the code output by
our compiler (C++ code in our implementation). It computes over the
public parts of the program, emitting a circuit that can later be
evaluated using a cryptographic MPC protocol. The semantics also
\emph{flattens} the arrays so that the circuits are unaware of the
array structure. Crucially, this phase of the semantics does not have
access to the secrets.

Figure~\ref{fig:tgtruntime} shows the runtime syntax. A wire id range
$r$ denotes a set of circuit wires that carry the value of a secret
shared variable in the output circuit (we will concretely define these
values later as part of the circuit semantics). Circuit gates $\kappa$
are $r$, input gates, \kw{add}, \kw{gt}, and \kw{mux} gates, and
coerce gates $\widetilde{w} \rhd m$. Target values $\widetilde{v}$
then consist of base values $c$ and $\kappa$, and arrays of base
values.

Figure~\ref{fig:tgtsem} shows the judgments for the target
semantics. We first focus on the expression evaluation judgment
$\widetilde{\rho} \vdash \widetilde{e} \Downarrow
\widetilde{v}$. Rules ({\sc{S-PAdd}}) and ({\sc{S-SAdd}}) illustrate
the significance of the operator labels. In particular, the rule
({\sc{S-PAdd}}) evaluates a public addition $\widetilde{e_{1}}
+_{\mathcal{P}} \widetilde{e_{2}}$ to $n_{1} + n_{2}$, similar to the
source language. In contrast, the rule
({\sc{S-SAdd}}) evaluates a secret addition $\widetilde{e_{1}}
+_{\mathcal{A}} \widetilde{e_{2}}$ to an arithmetic add
gate $\kw{add}\:\kappa_{1}\:\kappa_{2}$. Foreshadowing on the
compilation rules to be presented later, the target
expressions from our compiler never have $\widetilde{e_{1}}
+_{\mathcal{B}} \widetilde{e_{2}}$, as our compiler is aware that $+$
is more performant using arithmetic circuit, rather than a boolean
one. Rules ({\sc{S-PCond}}) and ({\sc{S-SCond}})
are along the similar lines. Rule ({\sc{S-PCond}}) evaluates a public
conditional to the value from one of the branches, while the rule
({\sc{S-SCond}}) evaluates to a \kw{mux} gate that takes input
circuits from the conditional ($\kappa_{1}$) and both the branches
($\kappa_{2}$ and $\kappa_{3}$). Once again for performance reasons,
the target expressions from our compiler do not have
$\lcond{\mathcal{A}}{e_{1}}{e_{2}}{e_{3}}$. Rules
({\sc{S-Coerce}}) and ({\sc{S-Inp}}) evaluate to the coerce and input
gates respectively.

Statement evaluation $\widetilde{\rho} \vdash \widetilde{s} \Downarrow
\widetilde{\rho}_{1}; \chi$ evaluates a target statement to produce a new
environment $\widetilde{\rho}_{1}$, and a circuit $\chi$. $\chi$ (shown
in Figure~\ref{fig:tgtlang}) is either empty, \kw{bind}-ing of a
circuit gate $\kappa$ to range $r$, \kw{out} gate, or a sequence of
circuits. Rules ({\sc{S-DeclC}}), ({\sc{S-DeclCkt}}), and
({\sc{S-DeclCktA}}) show the variable declaration cases. Rule
({\sc{S-DeclC}}) shows the case when $\widetilde{e}$ evaluates to
$\widetilde{v}$, where $\widetilde{v}$ is either a constant or an
array of constants. In this case, the mapping $x \mapsto
\widetilde{v}$ is added to the environment, and the resulting circuit
is empty. When 
$\widetilde{e}$ evaluates to a circuit $\kappa$, rule
({\sc{S-DeclCkt}}) picks a fresh $r$, adds the mapping $x \mapsto r$
to the environment, and outputs the circuit
$\kw{bind}\:\kappa\:r$. Rule ({\sc{S-DeclCktA}}) is analogous for
$\widetilde{e}$ evaluating to an array of circuits. The variable
assignment rules (not shown in the figure) are similar. Rule
({\sc{S-WriteCkt}}) shows the case for writing to an array, where the
array contents are secret. Finally, rule ({\sc{S-Out}}) compiles to an
\kw{out} circuit.

@@

During the circuit evaluation, the wire ranges $r$ are
mapped to (random) strings $b$. The semantics of these strings is
given by pairs of encryption-decryption functions, written as
$\mathcal{E}_{m}$ and $\mathcal{D}_{m}$ (where $m$ is either
$\mathcal{A}$ or $\mathcal{B}$). More concretely,
$\mathcal{E}_{m}(c)$ returns a pair of two string $b_{1}$ and $b_{2}$
(shares of the two parties), with the property that
$\mathcal{D}_{m}(b_{1}, b_{2}) = c$. We assume that the underlying MPC
protocol instantiates $\mathcal{E}_{m}$ and $\mathcal{D}_{m}$
appropriately.

The circuit semantics is shown in Figure~\ref{fig:cktsem} using the
judgments $\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \kappa
\Downarrow b_{1}, b_{2}$, and $\widehat{\rho}_{1}, \widehat{\rho}_{2}
\vdash \chi \Downarrow \widehat{\rho}'_{1}, \widehat{\rho}'_{2}; O$,
where $\widehat{\rho}_{1}$ and $\widehat{\rho}_{2}$ are the circuit
environments of the two parties, mapping wire ranges $r$ to strings
$b$. Rule ({\sc{C-Add}}) evinces the pattern for evaluating circuit
gates $\kappa$. To evaluate $\kw{add}\:\kappa_{1}\:\kappa_{2}$, the
rule first evaluates $\kappa_{1}$ to $(b_{11}, b_{21})$ and
$\kappa_{2}$ to $(b_{12}, b_{22})$. Shares $(b_{11}, b_{21})$ are then
combined using $\mathcal{D_{\mathcal{A}}}$ to $n_{1}$, and similarly
$(b_{12}, b_{22})$ are combined to $n_{2}$. The final output of the
\kw{add} gate is then $\mathcal{E}_{\mathcal{A}}(n_{1} + n_{2})$. Note
that this is a functional description of how the \kw{add} gate
evaluates, of course, concretely $n_{1}$ and $n_{2}$ are not observed
by the parties. Rule ({\sc{C-Coerce}}) re-encrypts the shares using the
scheme for $m$. Coming to the evaluation of circuits, the evaluation
of \kw{bind} updates the mapping of $r$ in the input environments, and
the rule ({\sc{C-Out}}) outputs the clear value $c$ to the
observations.

@@

     \inferrule*[lab={\footnotesize{T-Cond}}]
               {
                 \Gamma \vdash e : \kw{bool}^{\ell} \leadsto \widetilde{e}\\\\
                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \sigma^{\ell_{1}} \leadsto \widetilde{e_{i}}\\\\
                 \ell = \mathcal{P} \vee (\ell = \mathcal{B} \wedge \ell' =\mathcal{B})
               }
               {
                 \Gamma \vdash \cond{e}{e_{1}}{e_{2}} : \sigma^{\ell_{1}} \leadsto \lcond{\ell}{\widetilde{e}}{\widetilde{e_{1}}}{\widetilde{e_{2}}}
               }

@@

  \end{array}
  \]
  \\\\
    \fbox{$\Gamma \vdash s \leadsto \widetilde{s} \mid \Gamma_{1}$}
  \[
  \\
  \begin{array}{c}
     \inferrule*[lab={\footnotesize{T-Decl}}]
               {
                 \psi = \sigma \Rightarrow \tau = \sigma^{\ell}\\\\
                 \psi = \sigma[n] \Rightarrow \tau = \sigma^{\ell}[n]\\\\
                 \Gamma \vdash e : \tau \leadsto \widetilde{e}
               }
               {
                 \Gamma \vdash \psi\:x = e \leadsto \tau\:x = \widetilde{e} \mid \Gamma, x:\tau
               }

@@

     \inferrule*[lab={\footnotesize{T-For}}]
               {
                 \Gamma_{1} = \Gamma, x:\kw{uint}^{\mathcal{P}}\\\\
                 \Gamma_{1} \vdash \loops{x}{n_{2}}{s} \leadsto \loops{x}{n_{2}}{\widetilde{s}} \mid \Gamma_{1}
               }
               {
                 \Gamma \vdash \for{x}{n_{1}}{n_{2}}{s} \leadsto \for{x}{n_{1}}{n_{2}}{\widetilde{s}} \mid \Gamma
               }

@@

\subsubsection*{Compilation judgments} We now come to the compilation
rules shown in Figure~\ref{fig:compile}. We present the rules in a
declarative style, where the rules are non-syntax
directed, and the labels $\ell$ are chosen
non-deterministically. Section~\ref{sec:impl} describes the inference
scheme in our implementation.
